---
title: "Gov 50 Final Project"
author: "Maya O'Dea"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to sports or fashion. Whether it be data relating to sports performance or fashion trends, both industries are intriguing and use data science in different ways. 

## Proposal

Does the valence of Taylor Swift's songs affect their dancability? In this study, I plan to examine the danceability score Spotify has given each of Taylor Swift's songs and how it relates to the valence score on a scale of 0 to 1. The valance score measures the musical positiveness conveyed by the track. High valence means a happier, more cheerful, and more euphoric sounds versus low valence which are more of sad, depressed, or angry sounds.





```{r}
taylor_swift <- read.csv("/Users/mayaodea/Downloads/archive (4)/taylor_all_songs.csv")

```

```{r}
library(tidyverse)
taylor_swift |>
  filter(album_name == "repuatation" | album_name == "1989" | album_name == "evermore" | album_name == "Speak Now" | album_name == "Fearless (Taylor's Version" | album_name == "folklore" | album_name == "Lover" | album_name == "Midnights" | album_name == "Red (Taylor's Version)" | album_name == "Taylor Swift") |>
  ggplot(mapping = aes(x = valence, y = danceability)) +
  geom_point() +
  facet_wrap(~ album_name)
```

```{r}
dance_vibe <- lm(danceability ~ valence, data = taylor_swift)
dance_vibe
```

In analysis of the scatter plots categorized by album, there seems to be a positive correlation between valence and danceability. A greater valance, or more positive sounding song, seems to elicit greater danceability, whereas a lower valence, more negative sounding song, means less danceability. When running a regression, the correlation coefficient comes out to about 0.2, which is slight but positive. 


















