---
title: "Gov 50 Final Project"
author: "Maya O'Dea"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction 

Does self-employment have an effect on total amount of sleep? Does an individual's age and hourly wage have an effect as well? In this study, I plan to examine whether there is a relationship between being self-employed and total minutes of sleep per week. I will also be investigating the effects of age and hourly wage to control for confounders. Self-employment means working for oneself rather than an employer, many times running a business or a partner in trade. It can be considered a very taxing and inconsistent job, while also having the upside of managing yourself and your own hours. The trade-offs of being self-employed make it an interesting topic to look into and analyze its effects on one's lifestyle. Sleep, in particular, is a very important part of health, significant in maintaining several functions of the brain and body. I hypothesize those who are self-employed get larger amounts of sleep per week because of their greater freedom and flexibility. Those who are employed by an employer usually work a set of hours each day or week, giving them less time to balance the demands of work and life, let alone sleep. I also surmise age will have a confounding effect, but hourly wage will not.      

## Data Section

The data set used in this study focuses on sleep patterns, containing various characteristics and indicators on a group of individuals ages 23 to 65. From the data science platform Kaggle, this data is a cross-sectional design, comparing an individual's total number of sleep for one week. The data was collected via expert coding and contains information on wages, marital status, and education besides self-employment status and sleep per week.

Self-employment is the independent, explanatory variable and sleep is the dependent, outcome variable. Self-employment is coded as 1 if the individual is self-employed and 0 if not. The week long total minutes of sleep includes both sleep at night and naps at any point in the day. The histogram below visualizes the distribution of total minutes of sleep per week. The blue vertical line intercepts with a mean around 3,400 minutes a week of sleep, or 56 hours a week, averaging 8 hours a night. Age is measured in years and hourly wage is measured in dollars.          

```{r}
library(tidyverse)
sleep <- read_csv("~/Downloads/sleep75.csv")

sleep_count <- sleep |> 
  ggplot(mapping = aes(x = slpnaps)) +
  geom_histogram() +
  labs(x = "Weekly Total Minutes of Sleep",
       y = "Count",
       title = "Distribution of Sleep") +
  geom_vline(aes(xintercept = mean(slpnaps)), color = "dodgerblue") 
sleep_count

avgsleep <- sleep |> summarize(Average_sleep_per_week = mean(slpnaps))
knitr::kable(avgsleep)
```

## Results 

After analyzing the relationship between between total minutes of sleep a week and being self-employed or not, an average treatment effect of about -54.07 was found. This means those who are self-employed are more likely to get less sleep than those employer-employed by 54 minutes. This is not a huge difference, but it is about an hour less sleep a week. The boxplot visualizes this slight difference, with those being employer-employed having a greater number of individuals getting more sleep than those self-employed. However, those employer-employed do have a greater spread and greater number of outliers in the data, signifying more variance in total minutes of sleep per week. The 95% confidence interval found was between -165.56 to 54.13 which proves to be a wide range, containing the average treatment effect found. This range means that 95% of the time the truth will be found in this interval. So, the results stated above may need to be taken with a grain of salt due to the wide range of values.

When visualizing the first confounder of hourly wages, there seems to be a negative relationship found. After running the regression, it seems that as hourly wages increase, total minutes of sleep a week decreases. This was quantified in the -11.9 correlation coefficient, which was statistically significant considering an alpha level of 0.05. The coloring by employer on the scatterplot also supports the ideas concluded above between self-employed and employer-employed. One may be able to see a greater number of employer-employed individuals getting more sleep, however there is a large and varying range. When investigating the second confounder of age, a correlation coefficient of about 5.09 was found. The greater an individual's age, the more minutes of sleep a week they get, found statistically significant on an alpha level of 0.05. Lastly, when running the multiple regression with all three variables, each variable's correlation coefficient decreased. Both confounder variables' p-values showed statistical significance on an alpha level of 0.05, however the self-employment indicator did not exemplify significance on this alpha level. Even so, the adjusted R-squared lowered slightly by 0.004, meaning the added covariates of age and hourly wage do not help predict total minutes of sleep, not being confounders. This is due to the fact that when R-squared values decrease, the added variables do not add value to the model.



# Employer and Sleep Average Treatment Effect & Barplot
```{r}
library(infer)
sleep_selfe_ate <- sleep |>
  group_by(selfe) |>
  summarize(mean_slpnaps = mean(slpnaps)) |>
  pivot_wider(names_from = selfe,
              values_from = mean_slpnaps) |>
  mutate(ATE = `1` - `0`)
knitr::kable(sleep_selfe_ate)

sleep <- sleep |>
  mutate(employ = if_else(selfe == 1, "Self Employed", "Employer Employed"))

barplot_sleep <- sleep |>
  ggplot(mapping = aes(x = employ, y = slpnaps)) +
  geom_boxplot() +
  labs(x = "Employer",
       y = "Total Minutes of Sleep a Week", 
       title = "Employer and Sleep")
barplot_sleep
  
```

# Employer and Sleep Confidence Intervals
```{r}
sleep_dist <- sleep |>
  specify(slpnaps ~ employ) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "diff in means", order = c("Self Employed", "Employer Employed"))

knitr::kable(get_confidence_interval(sleep_dist, level = 0.95))

```

# Wages & Sleep Scatter Plot 
```{r}
scatter_sleep <- sleep |>
  ggplot(mapping = aes(x = hrwage, y = slpnaps, color = employ)) +
  geom_point(alpha = 0.5) +
  labs(x = "Hourly Wage",
       y = "Total Minutes of Sleep a Week",
       color = "Employer",
       title = "Wages and Sleep by Employer")
scatter_sleep
````
# Hourly Wage & Sleep Regression 
```{r}
slpwage_regression <- lm(slpnaps ~ hrwage, data = sleep)


library(broom)
tidy(slpwage_regression) |>
  knitr::kable(digits = 3)

````
# Age & Sleep Regression
```{r}
slpage_reg <- lm(slpnaps ~ age, data = sleep)
tidy(slpage_reg) |>
  knitr::kable(digits = 3)
```


# Multiple Regression
```{r}
mult_reg <- lm(slpnaps ~ selfe + hrwage + age, data = sleep)
mult_reg

modelsummary::modelsummary(list(slpwage_regression, slpage_reg, mult_reg))
```

## Conclusion

Self-employment has a slight negative effect on total amount of sleep per week, controlling both hourly wage and sleep. Both added variables did not have confounding effects. These conclusions disprove my original hypothesis about individuals who are self-employed getting a greater amount of sleep as well as age being a confounder. My hypothesis on hourly wage not being a confounder was correct. Nonetheless, my analysis may be threatened by the statistical insignificance of the self-employment indicator's correlation coefficient on the multiple regression. The large range of the confidence interval found also poses an unsureness to my analysis. In addition about 170 observations were deleted due to missingness in the multiple regression, limiting the data as well. In the future, the analysis may be improved with a greater number of participants and more in depth analysis on an individual's type of job. Having more individuals that are self-employed may provide a better gauge on the analysis. Sleep quality also may be investigated with more time and money available.


















